---
title: "STAT3006 A3 T3 KNN"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
<br>

```{r}
train <- read.csv(file = 'mnist_train.csv')
test <- read.csv(file = 'mnist_test.csv')
```

##### a)
```{r}
dim(train)
dim(test)
table(train$label)
table(test$label)

x <- as.matrix(train[,2:785],60000,784)
library(ggplot2)
library(matrixStats)
sds <- colSds(x)
qplot(sds,bins=256)
```

```{r}
# Prepare data
x_train <- train[,2:785] # train without label
y_train <- factor(train[,1]) # train label
x_test <- test[,2:785] # test without label
y_test <- factor(test[,1]) # test label

# remove the zero variance columns for training set
library(caret)
x_nzv <- nearZeroVar(x_train)
x_col <- setdiff(1:784, x_nzv)
x_train_nzv <- x_train[,x_col]

```


##### b)

```{r}
# improve efficiency
library(doParallel)
registerDoParallel(cores=8)

# train knn model
library(caret)
set.seed(123)
ctrl <- trainControl(method = "cv", number=5, p=.9) # 5-fold CV
knnFit <- train(x_train_nzv, y_train, method = "knn", trControl = ctrl)
plot(knnFit) # choose k value

```

Predictions for test set:
```{r}
# predictions for test set
pred_knn <- predict(knnFit, x_test)
conf_knn_test <- as.matrix(confusionMatrix(pred_knn, y_test))
conf_knn_test

```

Error rates for test set:
```{r}
# Error rates for test set
digit = c(0:9)
test_error <- data.frame(digit = c(digit,"overall"), error_rate=NA, CI_left=NA, CI_right=NA)

# class-specific error rates & CI
for (n in digit){
  all = sum(conf_knn_test[,n+1])
  success = conf_knn_test[n+1,n+1]
  fail = all - success
  CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
  # record error rate & CI
  test_error$error_rate[test_error$digit==n] <- fail/all
  test_error$CI_left[test_error$digit==n] <- CI[1]
  test_error$CI_right[test_error$digit==n] <- CI[2]
}

# overall error rate & CI
all = sum(conf_knn_test)
success = sum(diag(conf_knn_test))
fail = all - success
success = sum(diag(conf_knn_test))
CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
test_error$error_rate[test_error$digit=='overall'] <- fail/all
test_error$CI_left[test_error$digit=='overall'] <- CI[1]
test_error$CI_right[test_error$digit=='overall'] <- CI[2]

test_error
```

##### c) 
```{r}
# predictions for training set
pred_knn_train <- predict(knnFit, x_train) # predict test set
conf_knn_train <- as.matrix(confusionMatrix(pred_knn_train, y_train))
conf_knn_train
```

Error rates for training set
```{r}
# error rates for training set
train_error <- data.frame(digit = c(digit,"overall"), error_rate=NA, CI_left=NA, CI_right=NA)

# class-specific error rates & CI
for (n in digit){
  all = sum(conf_knn_train[,n+1])
  success = conf_knn_train[n+1,n+1]
  fail = all - success
  CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
  # record error rate & CI
  train_error$error_rate[train_error$digit==n] <- fail/all
  train_error$CI_left[train_error$digit==n] <- CI[1]
  train_error$CI_right[train_error$digit==n] <- CI[2]
}

# overall error rate & CI
all = sum(conf_knn_train)
success = sum(diag(conf_knn_train))
fail = all - success
CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
train_error$error_rate[train_error$digit=='overall'] <- fail/all
train_error$CI_left[train_error$digit=='overall'] <- CI[1]
train_error$CI_right[train_error$digit=='overall'] <- CI[2]

train_error
```



##### e)
```{r}
# Visualize data
library(RColorBrewer)
BNW <- c("white", "black")
CUSTOM_BNW <- colorRampPalette(colors = BNW)
rotate <- function(x) t(apply(x, 2, rev))

```

Successful examples of each digit: 
```{r}
# Successful examples of each digit from test set
par(mfrow = c(3, 4), mar=c(2, 2, 2, 2), pty ='s')
for (n in c(0:9)){
  knn_digit <- which(pred_knn == y_test & y_test == n) # correctly classified
  index = sample(knn_digit,1)
  m = rotate(matrix(unlist(x_test[index,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[index], "Predicted:", pred_knn[index]))}

```


##### f)
Failed examples of each digit from train set: 
```{r}
# failed examples of each digit from train set
par(mfrow = c(3, 4), mar=c(2, 2, 2, 2), pty ='s')
for (n in c(0:9)){
  knn_digit <- which(pred_knn_train != y_train & y_train == n) # misclassification
  index = sample(knn_digit,1)
  m = rotate(matrix(unlist(x_train[index,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[index], "Predicted:", pred_knn_train[index]))}

```

#### g)
```{r}
# successful & failed examples of 1
par(mfrow = c(2, 4), mar=c(2, 2, 2, 2), pty ='s')

knn_digit_s1 <- which(pred_knn == y_test & y_test== 1) # successful example of 1
index = sample(knn_digit_s1,4)
for (i in index){
  m = rotate(matrix(unlist(x_test[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[i], "Predicted:", pred_knn[i]))
}
knn_digit_f1 <- which(pred_knn_train != y_train & y_train== 1) # failed example of 1
index = sample(knn_digit_f1,4)
for (i in index){
  m = rotate(matrix(unlist(x_train[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[i], "Predicted:", pred_knn_train[i]))
}

```

```{r}
# successful & failed examples of 7
par(mfrow = c(2, 4), mar=c(2, 2, 2, 2), pty ='s')

knn_digit_s1 <- which(pred_knn == y_test & y_test== 7) # successful example of 7
index = sample(knn_digit_s1,4)
for (i in index){
  m = rotate(matrix(unlist(x_test[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[i], "Predicted:", pred_knn[i]))
}
knn_digit_f1 <- which(pred_knn_train != y_train & y_train== 7) # failed example of 7
index = sample(knn_digit_f1,4)
for (i in index){
  m = rotate(matrix(unlist(x_train[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[i], "Predicted:", pred_knn_train[i]))
}

```

## SVM

#### b) error rates for test set
```{r}
# improve efficiency
library(doParallel)
registerDoParallel(cores=8)

# train SVM model
library(caret)
ctrl <- trainControl(method = "cv", number = 5, verboseIter = F, allowParallel = T) # 5-fold CV
SVMfit <- train(x_train, y_train, method = "svmRadial", trControl = ctrl)
SVMfit # choose sigma and cost value

```

Predictions for test set:
```{r}
# predictions for test set
pred_SVM <- as.numeric(predict(SVMfit, x_test))-1
conf_SVM_test <- as.matrix(confusionMatrix(pred_SVM), y_test)
conf_SVM_test

```

Error Rates for test set:
```{r}
# error Rates for test set
test_error <- data.frame(digit = c(digit,"overall"), error_rate=NA, CI_left=NA, CI_right=NA)

# class-specific error rates & CI
for (n in digit){
  all = sum(conf_SVM_test[,n+1])
  success = conf_SVM_test[n+1,n+1]
  fail = all - success
  CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
  # record error rate & CI
  test_error$error_rate[test_error$digit==n] <- fail/all
  test_error$CI_left[test_error$digit==n] <- CI[1]
  test_error$CI_right[test_error$digit==n] <- CI[2]
}

# overall error rate & CI
all = sum(conf_SVM_test)
success = sum(diag(conf_SVM_test))
fail = all - success
CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
test_error$error_rate[test_error$digit=='overall'] <- fail/all
test_error$CI_left[test_error$digit=='overall'] <- CI[1]
test_error$CI_right[test_error$digit=='overall'] <- CI[2]

test_error

```



#### c) error rates for training set
```{r}
# predictions for training set
pred_SVM_train <- as.numeric(predict(SVMfit, x_train))-1
conf_SVM_train <- as.matrix(confusionMatrix(factor(pred_SVM_train), y_train))
conf_SVM_train

```

Error rates for training set:
```{r}
# error rates for training set
train_error <- data.frame(digit = c(digit,"overall"), error_rate=NA, CI_left=NA, CI_right=NA)

# class-specific error rates & CI
for (n in digit){
  all = sum(conf_SVM_train[,n+1])
  success = conf_SVM_train[n+1,n+1]
  fail = all - success
  CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
  # record error rate & CI
  train_error$error_rate[train_error$digit==n] <- fail/all
  train_error$CI_left[train_error$digit==n] <- CI[1]
  train_error$CI_right[train_error$digit==n] <- CI[2]
}

# overall error rate & CI
all = sum(conf_SVM_train)
success = sum(diag(conf_SVM_train))
fail = all - success
CI <- binom.test(c(fail,success), conf.level = 0.95)$conf.int
train_error$error_rate[train_error$digit=='overall'] <- fail/all
train_error$CI_left[train_error$digit=='overall'] <- CI[1]
train_error$CI_right[train_error$digit=='overall'] <- CI[2]

train_error

```

##### e)
```{r}
# Visualize data
library(RColorBrewer)
BNW <- c("white", "black")
CUSTOM_BNW <- colorRampPalette(colors = BNW)
rotate <- function(x) t(apply(x, 2, rev))
```

Successful examples of each digit: 
```{r}
# successful examples of each digit
par(mfrow = c(3, 4), mar=c(2, 2, 2, 2), pty ='s')
for (n in c(0:9)){
  SVM_digit <- which(pred_SVM == y_test & y_test == n) # correctly classified
  index = sample(SVM_digit,1)
  m = rotate(matrix(unlist(x_test[index,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[index], "Predicted:", pred_SVM[index]))}
```

##### f)
Failed examples of each digit from train set: 
```{r}
# failed examples of each digit from train set
par(mfrow = c(3, 4), mar=c(2, 2, 2, 2), pty ='s')
for (n in c(0:9)){
  SVM_digit <- which(pred_SVM_train != y_train & y_train == n) # misclassification
  index = sample(SVM_digit,1)
  m = rotate(matrix(unlist(x_train[index,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[index], "Predicted:", pred_SVM_train[index]))}

```

#### g)
```{r}
# successful & failed examples of 1
par(mfrow = c(2, 4), mar=c(2, 2, 2, 2), pty ='s')

SVM_digit_s1 <- which(pred_SVM == y_test & y_test== 1) # successful example of 1
index = sample(SVM_digit_s1,4)
for (i in index){
  m = rotate(matrix(unlist(x_test[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[i], "Predicted:", pred_SVM[i]))
}
SVM_digit_f1 <- which(pred_SVM_train != y_train & y_train== 1) # failed example of 1
index = sample(SVM_digit_f1,4)
for (i in index){
  m = rotate(matrix(unlist(x_train[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[i], "Predicted:", pred_SVM_train[i]))
}

```

```{r}
# successful & failed examples of 7
par(mfrow = c(2, 4), mar=c(2, 2, 2, 2), pty ='s')

SVM_digit_s1 <- which(pred_SVM == y_test & y_test== 7) # successful example of 7
index = sample(SVM_digit_s1,4)
for (i in index){
  m = rotate(matrix(unlist(x_test[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_test[i], "Predicted:", pred_SVM[i]))
}
SVM_digit_f1 <- which(pred_SVM_train != y_train & y_train== 7) # failed example of 7
index = sample(SVM_digit_f1,4)
for (i in index){
  m = rotate(matrix(unlist(x_train[i,-1]),ncol = 28,byrow = T))
  image(m, col=CUSTOM_BNW(255), main = paste("Actual:", y_train[i], "Predicted:", pred_SVM_train[i]))
}
```

